Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-03-17 ident_manu.Rnw
areal views to plot

% R dependencies
<<include = F, cache = F>>=
source('R/funcs.r')
library(tidyr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(Hmisc)
library(ggrepel)
library(english)
library(purrr)
@

@@

>>>>

% R dependencies
<<include = F, cache = F>>=
source('R/funcs.r')
library(tidyr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(Hmisc)
library(ggrepel)
library(english)
library(purrr)
library(grid)
library(lubridate)
# library(raster)
library(rgeos)
library(ggpolypath)
library(readxl)
library(ggmap)
@

@@

<<areachg, eval = T, fig.cap = 'Estimated hypoxic area and monthly averages (km$^2 \\times$ 1000) using the default parameter settings (a, b) and the differences using a 50\\% increase in each parameter (c, d).  The selected parameters were the top sensitive in each parameter category for \\ac{do} in \\cref{fig:sensalltile,tab:dosens}. Error bars are 95\\% confidence intervals for the monthly averages in (b) and the difference in monthly averages (d) from the perturbed and default results.  Areal extent of estimated hypoxia on August 29\\textsuperscript{th} using the defaulat parameters (grey) and estimated changes with parameter perturbations is shown in (e).', fig.height = 7, fig.width = 7, echo = F, cache = TRUE, out.width = "\\textwidth">>=
# ##
# # area time series
# areats <- read_excel('ignore/runcompare.xlsx', sheet = 'area')
# 
# ##
# # get all arrays of hypox area for each parameter
# shts <- excel_sheets('ignore/runcompare.xlsx') %>% 
#   grep('array', ., value = T)
# latlon <- read_excel('ignore/runcompare.xlsx', sheet = 'Latlon') %>% 
#   list(
#     lat = c(.[,1])[[1]], 
#     lon = as.numeric(names(.)[-1])
#     ) %>% 
#   .[c('lat', 'lon')]
# 
# ls <- vector('list', length = length(shts))
# names(ls) <- shts
# for(sht in shts){
#  
#   cat(sht, '\t')
#   
#   # read the sheet, interpolate for higher res
#   tmp <- read_excel('ignore/runcompare.xlsx', sheet = sht, na = '-9999', col_names = F) 
#   tmp <- tmp[1:176,1:320] %>% 
#     as.matrix
#   obs <- list(x = latlon$lat, y = latlon$lon, z = tmp)
#   newgrd <- expand.grid(
#       x = approx(latlon$lat, n = 4 * ncol(tmp))$y,
#       y = approx(latlon$lon, n = 4 * nrow(tmp))$y
#       )
#   int_val <- fields::interp.surface(
#     obj = obs,
#     loc = newgrd
#     )
#   tmp <- data.frame(newgrd, tmp = int_val) %>%
#     spread(x, tmp) %>%
#     dplyr::select(-y) %>%
#     as.matrix %>% 
#     .[, ncol(.):1] %>% 
#     t
#   
#   # subset hypoxic areas, make polygon
#   # get only the large areas
#   tmp[tmp >= 64] <- NA
#   tmp[!is.na(tmp)] <- 1
#   tmp <- raster(tmp, 
#     xmn = min(latlon$lon), 
#     xmx = max(latlon$lon),   
#     ymn = min(latlon$lat), 
#     ymx = max(latlon$lat)
#     )
#   tmp <- rasterToPolygons(tmp, dissolve = T) %>% 
#     disaggregate
#   areas <- transform(data.frame(tmp), area = gArea(tmp, byid = T))
#   filt <- with(areas, which(area > 0.12)) #quantile(area, 0.99)))
#   tmp <- tmp[filt, ] %>% 
#     fortify
#   
#   ls[[sht]] <- data.frame(parm = sht, tmp)
#    
# }
# 
# # combine output from each sheet
# arrs <- do.call('rbind', ls) %>% 
#   mutate(
#     parm = gsub('_array$', '', parm)
#     ) %>% 
#   # filter(lat > 28) %>% 
#   tibble::remove_rownames(.) %>% 
#   unite('grp', parm, group, remove = F)
# 
# # save
# save(areats, arrs, file = 'data/area.RData', compress = 'xz')

# area time series and 2d arrays (areats, arrs)
data(area)

# theme
mytheme <- theme_minimal() + 
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.title.y = element_blank(),
    axis.title.x = element_blank(), 
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"),
    legend.title = element_blank(),
    legend.position = 'top', 
    legend.text.align = 0
  )

# colors
cols <- RColorBrewer::brewer.pal(5, 'Spectral')

# format area time series 
areats <- areats %>% 
  dplyr::select(dts = 1, everything()) %>% 
  mutate(
    dts = as.Date(dts), 
    mo = month(dts, label = T)
  ) %>% 
  gather('var', 'val', -dts, -mo) %>% 
  mutate(
    var = gsub('\\+50%$', '', var),
    val = val / 1000
    ) %>% 
  spread(var, val) %>% 
  filter(mo %in% c('May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct'))

# formatted levels in order
levs <- c('umax', 'ZKa', 'astar490', 'KG1', 'Tref')
labs <- paste0(c('umax', 'ZKa', 'astar490', 'KG1', 'Tref(nospA+nospZ)'), '_1')
labs <- par_txt(labs, frm = 'exp')

##
# data for each plot

toplo1 <- areats

toplo2 <- gather(areats, 'var', 'val', -dts, -Original, -mo) %>% 
  mutate(
    valper = val - Original,
    var = factor(var, levels = levs, labels = labs)
  )

toplo3 <- areats %>%
  group_by(mo) %>% 
  summarise(
    valup = t.test(Original)$conf.int[2],
    valdn = t.test(Original)$conf.int[1],
    val = mean(Original)
  )

toplo4 <- gather(areats, 'var', 'val', -dts, -mo, -Original) %>% 
  group_by(mo, var) %>% 
  summarise(
    valup = t.test(val, Original)$conf.int[2],
    valdn = t.test(val, Original)$conf.int[1],
    val = mean(val) - mean(Original)
  ) %>% 
  ungroup %>% 
  mutate(
    var = factor(var, levels = levs, labels = labs)
  )

toplo5 <- filter(arrs, parm %in% 'original')

toplo6 <- filter(arrs, !parm %in% 'original') %>% 
  mutate(
    parm = factor(parm, levels = levs, labels = labs)
  )

##
# subplots

p1 <- ggplot(toplo1, aes(x = dts, y = Original)) + 
  geom_polygon() + 
  mytheme + 
  scale_x_date(date_labels = '%b', date_breaks = '1 month') +
  ggtitle(label = NULL, subtitle = '(a) Area with default parameters')

p2 <- ggplot(toplo2, aes(x = dts, y = valper, fill = var)) + 
  geom_ribbon(aes(ymin = 0, ymax = valper), colour = 'darkgrey', size = 0.35) + 
  mytheme %+replace%
  theme(legend.position = 'none') +
  scale_x_date(date_labels = '%b', date_breaks = '1 month') +
  scale_fill_manual(values = cols, labels = labs) +
  ggtitle(label = NULL, subtitle = '(c) Area differences with parameter perturbations')

p3 <- ggplot(toplo3, aes(x = mo, y = val)) + 
  geom_bar(stat = 'identity') + 
  geom_errorbar(aes(ymin = valdn, ymax = valup), width = 0.25, lwd = 0.35) + 
  mytheme +
  ggtitle(label = NULL, subtitle = '(b) Averages with default parameters')

pd <- position_dodge(0.75)
p4 <- ggplot(toplo4, aes(x = mo, y = val, fill = var)) + 
  geom_bar(stat = 'identity', position = pd, width = 0.75, colour = 'darkgrey', lwd = 0.35) + 
  geom_errorbar(aes(ymin = valdn, ymax = valup), position = pd, width = 0.25, size = 0.35) + 
  scale_fill_manual(values = cols, labels = labs) +
  mytheme +
  ggtitle(label = NULL, subtitle = '(d) Average differences with parameter perturbations')

pleg <- g_legend(p4)
p4 <- p4 + 
  mytheme  %+replace%
  theme(legend.position = 'none')

# basemap for area ts
ext <- make_bbox(arrs$long, arrs$lat, f = 0.01)
map <- get_stamenmap(ext, zoom = 9, maptype = "toner-lite")

p5 <- ggmap(map) + 
  geom_polypath(data = toplo5, aes(x = long, y = lat, group = grp), alpha = 0.5) +
  geom_path(data = toplo6, aes(x = long, y = lat, group = grp, colour = parm), alpha = 0.9, size = 1) + 
  scale_y_continuous(expand = c(0, 0)) + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_colour_manual(values = cols, labels = labs) +
  coord_equal() + 
  mytheme %+replace%
  theme(legend.position = 'none') +
  ggtitle(label = NULL, subtitle = '(e) Estimated areal extents on August 29')

pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
pC <- ggplot_gtable(ggplot_build(p3))
pD <- ggplot_gtable(ggplot_build(p4))
pE <- ggplot_gtable(ggplot_build(p5))
maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3])
pA$widths[2:3] <- maxWidth
pC$widths[2:3] <- maxWidth
pE$widths[2:3] <- maxWidth
maxWidth = unit.pmax(pC$widths[2:3], pD$widths[2:3], pE$widths[2:3])
pB$widths[2:3] <- maxWidth
pD$widths[2:3] <- maxWidth

ylab <- expression(paste('Hypoxic area ( ', km^2, ' x 1000)'))
grid.arrange(
  left = textGrob(ylab, rot = 90),
  arrangeGrob(
    arrangeGrob(pA, pC, ncol = 2),
    pleg, 
    arrangeGrob(pB, pD, ncol = 2), 
    pE,
    ncol = 1, 
    heights = c(1, 0.1, 1, 1.4)
  )
)

### 2017-03-17 ident_manu.tex
areal views to plot

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Thu Mar 9 13:12:38 2017 -0600, \href{https://github.com/fawda123/identifiability/commit/70c8da3c6d4aa0e50231198ae90027a250ce2ae5}{70c8da3c6d4aa0e50231198ae90027a250ce2ae5}}

@@

>>>>

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Sun Mar 12 12:53:46 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/39a5c73735c94c696b2422eddee88342245006dd}{39a5c73735c94c696b2422eddee88342245006dd}}

@@

\begin{figure}[!ht]

{\centering \includegraphics[width=\textwidth]{figs/areachg-1}

\caption{Estimated hypoxic area and monthly averages (km$^2 \times$ 1000) using the default parameter settings (a, b) and the differences using a 50\% increase in each parameter (c, d).  The selected parameters were the top sensitive in each parameter category for \ac{do} in \cref{fig:sensalltile,tab:dosens}. Error bars are 95\% confidence intervals for the monthly averages in (b) and the difference in monthly averages (d) from the perturbed and default results.  Areal extent of estimated hypoxia on August 29\textsuperscript{th} using the defaulat parameters (grey) and estimated changes with parameter perturbations is shown in (e).}\label{fig:areachg}
\end{figure}
